---
title: "Function to estimate power to detect significant heterogeneity in meta-analysis"
author: "Sara J. Weston"
date: "10/25/2019"
output: html_document
---

## Function

The following function was used to estimate the power of the significance test of heterogenity ($Q$) in a meta-analysis. Equations were based on [Valentine, Pigott, & Rothstein, 2010](https://journals.sagepub.com/doi/full/10.3102/1076998609346961).

Note that the tidyverse package is required.

```{r, results = 'hide', echo = F, warning = F, message = F}
library(tidyverse)
```


```{r}
Q.pwr = function(Q, nstudy, alpha = .05, plot = FALSE){

  require(tidyverse, attach.required = T, quietly = T, warn.conflicts = F)

  try(if(length(nstudy) > 1) stop("only one number can be entered for nstudy"))
  try(if(length(alpha) > 1) stop("only one number can be entered for alpha"))

  cv = qchisq(p = 1-alpha, df = nstudy-1, ncp = Q, lower.tail = T)
  typeII = pf(q = cv, df1 = cv, df2 = Q, lower.tail = F)
  power = 1 - typeII

  return.plot = NULL

  if(plot == TRUE){

    try(if(length(Q) < 2) warning("multiple values of Q needed for plot"))

    try(if(length(Q) > 1)
      return.plot = data.frame(Q = Q, power = power) %>%
        ggplot(aes(x = Q, y = power )) +
        geom_point(size = 2) +
        geom_line() +
        theme_bw() +
        ggtitle(paste("Power estimates of heterogeneity by Q for", nstudy, "studies")))

  }

  if(!is.null(return.plot)) return(list(power, return.plot))
  if(is.null(return.plot)) return(power)

}
```


## Example: single estimate

The function requires input of a Q statistic (hypothetical if pre-data collection, actual if post-hoc power analysis) and the number of studies. $\alpha$  is assumed to be .05.

```{r}
Q.pwr(Q = 10, nstudy = 4)
```


## Example: multiple estimates

Q can take the form of a vector of multiple values, to estimate power at different levels of heterogeneity. 

```{r}
Q.pwr(Q = 1:20, nstudy = 4)
```

In addition, you can ask the function to plot these estimates.

```{r}
Q.pwr(Q = 1:20, nstudy = 4, plot = TRUE)
```



